Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-11 inc/hardware/magnetic_lenses/maglens.tex
tweak assembly view of mag lens design schematic
citations are now preceeded by non-breaking space (~)

\begin{figure}
  \centering
  \input{polepiece-design}
  \caption[Design schematic of our magnetic lens with a pole-piece]{
    Side-view schematic of the magnetic lens with a pole-piece designed for the UEM at UIC.
    On the center axis is the inner drift tube, with pole-piece gap separation disk.
    The end pieces are shown top and bottom.
    One bridge piece is shown detached (left) and one attached (right).
    Note that this figure is not drawn to the scale of the lenses at UIC, but rather to match the general schematic in \ref{fig:polepiece-schematic}.
  }
  \label{fig:polepiece-design}
\end{figure}

@@

El-Kareh and El-Kareh provide an analysis for symmetric lenses with magnetic shielding and a pole-piece (see Ref. \cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
They define two parameters,
\begin{equation}
  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2
\end{equation}
and
\begin{equation} \label{eq:elkareh_k2}
  \beta \equiv k^2 V_r / (NI)^2 \,\text{,}
\end{equation}
where $N$ is the number of loops, $I$ is the current in a single loop, $B_0$ is the maximum axial magnetic field strength, $R$ is the bore radius and $V_r$ is the relativity corrected acceleration potential, given by
\begin{equation}
  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V ) \,\text{.}
\end{equation}
Using the provided tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.
Since the drift tube has an outer diameter $D=$0.5in and the pole-piece gap length is $S$=0.1in, the lens has a ratio $S/D=0.2$.
By their table 8.2a this results in $\beta=0.0146$.
Then, to create a focal length of 4in ($f/R=16$) table 8.13 in Ref. \cite{el-kareh_electron_1970} gives a value for $k^2 \approx 0.06$.
Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.

>>>>

\begin{figure}
  \centering
  \input{polepiece-design}
  \caption[Design schematic of our magnetic lens with a pole-piece]{
    Side-view schematic of the magnetic lens with a pole-piece designed for the UEM at UIC.
    On the center axis is the inner drift tube, with pole-piece gap separation disk.
    The end pieces are shown top and bottom, bridges are on the left and right.
    The left side shows a bridge piece and one end before assembly.
    Note that this figure is not drawn to the scale of the lenses at UIC, but rather to match the general schematic in \ref{fig:polepiece-schematic}.
  }
  \label{fig:polepiece-design}
\end{figure}

@@

El-Kareh and El-Kareh provide an analysis for symmetric lenses with magnetic shielding and a pole-piece (see Ref.~\cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
They define two parameters,
\begin{equation}
  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2
\end{equation}
and
\begin{equation} \label{eq:elkareh_k2}
  \beta \equiv k^2 V_r / (NI)^2 \,\text{,}
\end{equation}
where $N$ is the number of loops, $I$ is the current in a single loop, $B_0$ is the maximum axial magnetic field strength, $R$ is the bore radius and $V_r$ is the relativity corrected acceleration potential, given by
\begin{equation}
  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V ) \,\text{.}
\end{equation}
Using the provided tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.
Since the drift tube has an outer diameter $D=$0.5in and the pole-piece gap length is $S$=0.1in, the lens has a ratio $S/D=0.2$.
By their table 8.2a this results in $\beta=0.0146$.
Then, to create a focal length of 4in ($f/R=16$) table 8.13 in Ref.~\cite{el-kareh_electron_1970} gives a value for $k^2 \approx 0.06$.
Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.

### 2013-04-11 inc/introduction.tex
citations are now preceeded by non-breaking space (~)

The development of laboratory-scale instrumentation for the direct visualization of fundamental ultrafast (i.e., sub-nanosecond) events with nano-scale (even atomic) spatial resolution is a current active area of research \cite{king_ultrafast_2005}.
In many areas of physics, chemistry and materials science a space-time resolution in the sub-1nm$\cdot$ps range is ideally required to allow for the observation of the fastest structural events.
A promising approach is to employ ultrashort pulses of electrons generated by a laser-driven 10-100keV photoelectron gun to probe the dynamics of the specimen that has been perturbed by a synchronized laser pulse \cite{king_ultrafast_2005}.
Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED) \cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting \cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003}, chemical bond breaking \cite{zewail_laser_1988}, etc.
Direct time-resolved imaging of similar events can be accomplished using transmission electron microscopy (TEM).
To date, such dynamic transmission electron microscopes (DTEMs) \cite{bostanjoglo_tracing_1980,domer_high-speed_2003} have been developed for both single-shot imaging with $\sim$$10^{5}$nm$\cdot$ps space-time resolution (e.g., 30ns electron pulses and 5-10nm spatial resolution \cite{lagrange_single-shot_2006,armstrong_practical_2007}) and $\sim$1nm$\cdot$ps space-time resolution ($\sim$100fs with a few nanometer resolution\cite{park_direct_2009}) in the multi-shot data acquisition mode of operation.
Both of these instruments employ retro-fitted standard electron microscope columns and are limited in performance mainly by space-charge effects.

Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects \cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities \cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).

In this thesis, I will present an extension of the work of Michalik and Sipe \cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment, to include the influence of linear external forces due to electron optical elements.
The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
For completeness, in Section \ref{sec:ms_model} I will introduce, in MKS units, the essential features of the AG model.
In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic external forces.
In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
I would like to note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.

Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid \cite{luiten_how_2004}.
This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
Consequently, the AG approach is applicable to both the single electron per pulse limit \cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate \cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
%Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.

In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the behaviors of ultrafast electron pulses in a UEM column.
Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion \cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance \cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle).

>>>>

The development of laboratory-scale instrumentation for the direct visualization of fundamental ultrafast (i.e., sub-nanosecond) events with nano-scale (even atomic) spatial resolution is a current active area of research~\cite{king_ultrafast_2005}.
In many areas of physics, chemistry and materials science a space-time resolution in the sub-1nm$\cdot$ps range is ideally required to allow for the observation of the fastest structural events.
A promising approach is to employ ultrashort pulses of electrons generated by a laser-driven 10-100keV photoelectron gun to probe the dynamics of the specimen that has been perturbed by a synchronized laser pulse~\cite{king_ultrafast_2005}.
Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED)~\cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting~\cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003}, chemical bond breaking~\cite{zewail_laser_1988}, etc.
Direct time-resolved imaging of similar events can be accomplished using transmission electron microscopy (TEM).
To date, such dynamic transmission electron microscopes (DTEMs)~\cite{bostanjoglo_tracing_1980,domer_high-speed_2003} have been developed for both single-shot imaging with $\sim$$10^{5}$nm$\cdot$ps space-time resolution (e.g., 30ns electron pulses and 5-10nm spatial resolution~\cite{lagrange_single-shot_2006,armstrong_practical_2007}) and $\sim$1nm$\cdot$ps space-time resolution ($\sim$100fs with a few nanometer resolution~\cite{park_direct_2009}) in the multi-shot data acquisition mode of operation.
Both of these instruments employ retro-fitted standard electron microscope columns and are limited in performance mainly by space-charge effects.

Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects~\cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).

In this thesis, I will present an extension of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment, to include the influence of linear external forces due to electron optical elements.
The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
For completeness, in Section \ref{sec:ms_model} I will introduce, in MKS units, the essential features of the AG model.
In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic external forces.
In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
I would like to note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.

Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations~\cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
Consequently, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
%Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.

In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the behaviors of ultrafast electron pulses in a UEM column.
Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle).

### 2013-04-11 inc/model_results/compound/compound.tex
killing TODOs
citations are now preceeded by non-breaking space (~)

Through the use of multiple electron optical components such as magnetic lenses and RF cavities and accelerators (properly region-of-influence limited by application of \ref{eq:reg_of_influence}), the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
However, for a nanosecond DTEM \cite{lagrange_nanosecond_2008,reed_evolution_2009}, the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs\cite{park_direct_2009} and UED experiments \cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).

@@

\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements \cite{armstrong_practical_2007}.
The displayed dynamics are for an electron pulse generated by a $\tau = $ 100fs $w = $ 0.1mm laser, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
In this simple system, an optimum RF cavity ($ \Omega = \pi v_{{ \scriptscriptstyle 0}} / d $) is surrounded by two magnetic lenses (positioned a realistic 5cm to either side of the RF cavity) that compensate for the electron beam divergence due to the non-zero $ \eta_{\smallT} (0) $ and the negative spatial lensing of the RF pulse compression cavity.
For the transverse dimension, this arrangement of electron optical elements is similar to the Cooke triplet in optics.
The magnetic lenses are assumed to be identical and wired in series, and thus have equal magnetic field strengths.
The three lensing elements are positioned at a distance of 15cm from the cathode to the front of the first magnetic lens and are adjusted so as to space-time image the electron pulse 5cm behind the second magnetic lens.
The dependencies of the the spatial variances, $\sigma_{\alpha}$, shown in \ref{fig:compound} as the electron pulse propagates through the multi-element system, clearly illustrate the effect of the magnetic lenses and RF cavity in the transverse ($\smallT$) and longitudinal ($z$) dimensions.
As the image-to-object distance ratio is less than unity in this case, both the transverse and longitudinal imaged pulse dimensions are reduced relative to their initial values. %TODO comment futher

As seen in Section \ref{sec:free_spacecharge}, with increasing pulse charge, $ N e $, space-charge effects in general will act to degrade the space-time imaging by increasing the three-dimensional pulse size delivered to the reference plane positioned 5cm behind the second magnetic lens.
At $ N = 10^{6} $, space-charge effects moderate the image demagnification due to the $\sim$2 object-to-image distance ratio in \ref{fig:compound}. 
In other words, for the displayed compound element system, which could be employed in UED experiments \cite{oudheusden_electron_2007}, the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
For a UEM, on the other hand, it may be necessary to separate longitudinally the space and time images produced by the compound element column.
This is because the inclusion of the microscope's objective lens will likely require the pulse's time focus to be at the objective's back focal plane (i.e., in the specimen) while the column's spatial image may need to be located at the objective's front focal plane (i.e., entrance aperture) for efficient beam transfer with minimal spherical aberration.
Fortunately, a compound component system of the type shown in \ref{fig:compound} with a RF cavity can meet this challenge since the longitudinal positions of the two foci can be manipulated independently.

>>>>

Through the use of multiple electron optical components such as magnetic lenses and RF cavities and accelerators (properly region-of-influence limited by application of \ref{eq:reg_of_influence}), the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
However, for a nanosecond DTEM~\cite{lagrange_nanosecond_2008,reed_evolution_2009}, the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs~\cite{park_direct_2009} and UED experiments~\cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).

@@

\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements~\cite{armstrong_practical_2007}.
The displayed dynamics are for an electron pulse generated by a $\tau = $ 100fs $w = $ 0.1mm laser, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
In this simple system, an optimum RF cavity ($ \Omega = \pi v_{{ \scriptscriptstyle 0}} / d $) is surrounded by two magnetic lenses (positioned a realistic 5cm to either side of the RF cavity) that compensate for the electron beam divergence due to the non-zero $ \eta_{\smallT} (0) $ and the negative spatial lensing of the RF pulse compression cavity.
For the transverse dimension, this arrangement of electron optical elements is similar to the Cooke triplet in optics.
The magnetic lenses are assumed to be identical and wired in series, and thus have equal magnetic field strengths.
The three lensing elements are positioned at a distance of 15cm from the cathode to the front of the first magnetic lens and are adjusted so as to space-time image the electron pulse 5cm behind the second magnetic lens.
The dependencies of the the spatial variances, $\sigma_{\alpha}$, shown in \ref{fig:compound} as the electron pulse propagates through the multi-element system, clearly illustrate the effect of the magnetic lenses and RF cavity in the transverse ($\smallT$) and longitudinal ($z$) dimensions.
As the image-to-object distance ratio is less than unity in this case, both the transverse and longitudinal imaged pulse dimensions are reduced relative to their initial values.

As seen in Section \ref{sec:free_spacecharge}, with increasing pulse charge, $ N e $, space-charge effects in general will act to degrade the space-time imaging by increasing the three-dimensional pulse size delivered to the reference plane positioned 5cm behind the second magnetic lens.
At $ N = 10^{6} $, space-charge effects moderate the image demagnification due to the $\sim$2 object-to-image distance ratio in \ref{fig:compound}. 
In other words, for the displayed compound element system, which could be employed in UED experiments~\cite{oudheusden_electron_2007}, the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
For a UEM, on the other hand, it may be necessary to separate longitudinally the space and time images produced by the compound element column.
This is because the inclusion of the microscope's objective lens will likely require the pulse's time focus to be at the objective's back focal plane (i.e., in the specimen) while the column's spatial image may need to be located at the objective's front focal plane (i.e., entrance aperture) for efficient beam transfer with minimal spherical aberration.
Fortunately, a compound component system of the type shown in \ref{fig:compound} with a RF cavity can meet this challenge since the longitudinal positions of the two foci can be manipulated independently.

### 2013-04-11 inc/model_results/lens_charge/lens_charge.tex
citations are now preceeded by non-breaking space (~)

In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses \cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry \cite{berger_dc_2009}.
The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

>>>>

In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses~\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

### 2013-04-11 inc/model_results/rf_cav/rf_cav.tex
killing TODOs
citations are now preceeded by non-breaking space (~)

In comparison to using magnetic lenses to compensate for transverse pulse broadening by using magnetic lenses, RF cavities are being employed in UED systems \cite{oudheusden_electron_2007}, and are expected to be employed in UEM systems as well, to compensate for longitudinal broadening in the ultrashort pulses.
\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and $N=10^5$ pulse charges.
The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and a region of influence super-Gaussian of order 2 ($n=2$ in \ref{eq:reg_of_influence}) to reflect the sharper ends of the RF cavity.
In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).

@@

%TODO why is this even worse than before?
The RF electric field amplitude $E_{0}$ of about 450kV/m required to achieve the temporal focusing 10cm behind the RF cavity in the dispersive limit is also consistent with expectations.
By equating the differential impulse applied by the RF cavity across the electron pulse duration to that required to compensate for dispersive pulse broadening before the RF cavity and then compress the pulse thereafter, one can show that the required RF field amplitude for a $\text{TM}_{010}$ cavity with $ d = \pi v_{\smallzero} / \Omega $ is given to a good approximation by
\begin{equation} \label{eq:RF_field_required}
  E_{0} \approx V_{DC} \left ( \frac{1}{ L_{1} } + \frac{1}{ L_{2} } \right ) \text{,}
\end{equation}
where $L_{1}$ is the distance from the pulse source (the DC photoelectron gun) to the RF cavity (the object distance) and $L_{2}$ is the distance to the temporal pulse focus behind the RF cavity (the image distance).
For the considered case, $V_{DC} = $ 20kV and $ L_{1} = L_{2} = $ 10cm, giving $ E_{0} \approx $ 400kV/m, which is in reasonably good agreement with the optimum value of 450kV/m employed for the simulation displayed in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}.
\ref{eq:RF_field_required} also clearly indicates that the DC photoelectron gun and $\text{TM}_{010}$ RF pulse compression cavity should be considered as a single system, since the temporal `focal length' of the RF cavity, $ V_{DC} / E_{0} $, is dependent upon the acceleration voltage.

The optimized RF cavity performance for the temporal compression of a pulse with the same initial photoemission conditions, but with $N = 10^{5}$ electrons, is shown in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1e5}.
In this case, intra-pulse space-charge effects act to further broaden the initial 100fs pulse duration to $\sim$6ps after the $\sim$1ns time-of-flight over the 10cm distance to the RF cavity.
The fact that almost the same RF field strength (495 kV/m) is required to generate a temporal focus at $z^{\prime} = $ 10cm is a direct consequence of the linear dependence of the longitudinal RF cavity force on $z$ (\ref{eq:RF_Force_z}), which means that the required restorative force is not dependent on $\sigma_{z}$ --- just as the focal length of a perfect optical lens is independent of its aperture size.
However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less ($\sim 48$ fs) than the original 100fs laser pulse duration.
This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression \cite{strickland_compression_1985}.
In this case, the $\sim$20-fold increase in the transverse pulse size (corresponding to a $\sim400\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.
Again, in analogy to optical imaging, a stronger (and suitably phased) RF field will result in an even shorter compressed pulse duration as the distance to the temporal focus behind the RF cavity is reduced.

>>>>

In comparison to using magnetic lenses to compensate for transverse pulse broadening by using magnetic lenses, RF cavities are being employed in UED systems~\cite{oudheusden_electron_2007}, and are expected to be employed in UEM systems as well, to compensate for longitudinal broadening in the ultrashort pulses.
\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and $N=10^5$ pulse charges.
The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and a region of influence super-Gaussian of order 2 ($n=2$ in \ref{eq:reg_of_influence}) to reflect the sharper ends of the RF cavity.
In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).

@@

The RF electric field amplitude $E_{0}$ of about 450kV/m required to achieve the temporal focusing 10cm behind the RF cavity in the dispersive limit is also consistent with expectations.
By equating the differential impulse applied by the RF cavity across the electron pulse duration to that required to compensate for dispersive pulse broadening before the RF cavity and then compress the pulse thereafter, one can show that the required RF field amplitude for a $\text{TM}_{010}$ cavity with $ d = \pi v_{\smallzero} / \Omega $ is given to a good approximation by
\begin{equation} \label{eq:RF_field_required}
  E_{0} \approx V_{DC} \left ( \frac{1}{ L_{1} } + \frac{1}{ L_{2} } \right ) \text{,}
\end{equation}
where $L_{1}$ is the distance from the pulse source (the DC photoelectron gun) to the RF cavity (the object distance) and $L_{2}$ is the distance to the temporal pulse focus behind the RF cavity (the image distance).
For the considered case, $V_{DC} = $ 20kV and $ L_{1} = L_{2} = $ 10cm, giving $ E_{0} \approx $ 400kV/m, which is in reasonably good agreement with the optimum value of 450kV/m employed for the simulation displayed in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}.
\ref{eq:RF_field_required} also clearly indicates that the DC photoelectron gun and $\text{TM}_{010}$ RF pulse compression cavity should be considered as a single system, since the temporal `focal length' of the RF cavity, $ V_{DC} / E_{0} $, is dependent upon the acceleration voltage.

The optimized RF cavity performance for the temporal compression of a pulse with the same initial photoemission conditions, but with $N = 10^{5}$ electrons, is shown in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1e5}.
In this case, intra-pulse space-charge effects act to further broaden the initial 100fs pulse duration to $\sim$6ps after the $\sim$1ns time-of-flight over the 10cm distance to the RF cavity.
The fact that almost the same RF field strength (495 kV/m) is required to generate a temporal focus at $z^{\prime} = $ 10cm is a direct consequence of the linear dependence of the longitudinal RF cavity force on $z$ (\ref{eq:RF_Force_z}), which means that the required restorative force is not dependent on $\sigma_{z}$ --- just as the focal length of a perfect optical lens is independent of its aperture size.
However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less ($\sim 48$ fs) than the original 100fs laser pulse duration.
This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression~\cite{strickland_compression_1985}.
In this case, the $\sim$20-fold increase in the transverse pulse size (corresponding to a $\sim400\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.
Again, in analogy to optical imaging, a stronger (and suitably phased) RF field will result in an even shorter compressed pulse duration as the distance to the temporal focus behind the RF cavity is reduced.
